---
title: "Simple and Mixed Model Analysis"
author: "Alexis Swanson"
date: "2/4/2021"
output: html_document
---

#Histograms of the 3 Approaches for tc
  
```{r, echo=FALSE, message=FALSE}

data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

hist(data$tc.B)
hist(data$tc.S)
hist(data$tc.I)

hist(log(data$tc.I))

```
The log of the Iterative Solution has the most normal distribution. We need to decide which solution to use - I am still not sure how to decide. 

I also need to finish the calibration and adjust the tc results based upon the updated coefficient. I will finish that this week.

# First Linear Model
  # All Predictors

```{r, echo=FALSE, message=FALSE}
data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

library(ggplot2)

model_simple <-lm(tc.I ~ Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_simple)

par(mfrow = c(2,2))
plot(model_simple)
par(mfrow = c(1,1))
qqnorm(resid(model_simple))
qqline(resid(model_simple))

boxplot(log(tc.I) ~ Test.Location, data=data)
boxplot(log(tc.I) ~ Land.Use, data=data)
boxplot(log(tc.I) ~ Soil.Texture, data=data)

```

# Linear Model with just EC and Moist
  # Based on the results those were the only significant factors
  
```{r, echo=FALSE, message=FALSE}
data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

model_simple2 <-lm(tc.I ~ EC + Cor.Moist, data = data)
summary(model_simple2)
par(mfrow = c(2,2))
plot(model_simple2)
par(mfrow = c(1,1))
qqnorm(resid(model_simple2))
qqline(resid(model_simple2))
```
  
# Linear Model excluding the outliers

```{r, echo=FALSE, message=FALSE}
data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

model_simple3 <-lm(tc.I ~ Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data[-c(18,23, 29, 38,39, 50, 26, 45, 33),])
summary(model_simple3)
par(mfrow = c(2,2))
plot(model_simple3)
par(mfrow = c(1,1))
qqnorm(resid(model_simple3))
qqline(resid(model_simple3))
```

# Linear Model excluding the outliers with significant factors
  # EC, Moist, D50 and Cu
  
```{r, echo=FALSE, message=FALSE}
data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

model_simple4 <-lm(tc.I ~ EC + Cor.Moist + D50 + Cu, data = data[-c(18,23, 29, 38,39, 50, 26, 45, 33),])
summary(model_simple4)
par(mfrow = c(2,2))
plot(model_simple4)
par(mfrow = c(1,1))
qqnorm(resid(model_simple4))
qqline(resid(model_simple4))
```

# Mixed Model with 2 Random Factors
  # Test Location and Land Use
  
```{r, echo=FALSE, message=FALSE}
data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

library(lme4)

model_mixed <-lmer(tc.I~(1|Test.Location) + (1|Land.Use) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed)
anova(model_mixed)
ranef(model_mixed)

par(mfrow = c(1,1))
plot(model_mixed)
qqnorm(resid(model_mixed))
qqline(resid(model_mixed))

```

# Mixed Model with only Test Location as Random Factor

```{r, echo=FALSE, message=FALSE}
data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

model_mixed2 <-lmer(tc.I~(1|Test.Location) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed2)
anova(model_mixed2)
ranef(model_mixed2)

par(mfrow = c(1,1))
plot(model_mixed2)
qqnorm(resid(model_mixed2))
qqline(resid(model_mixed2))

```

#Mixed Model with only Land Use as Random Factor

```{r, echo=FALSE, message=FALSE}
data <- read.csv("C:\\Users\\swans\\Documents\\Research\\Site_Data_Feb_2_2021.csv")

model_mixed3 <-lmer(tc.I~(1|Test.Location) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed3)
anova(model_mixed3)
ranef(model_mixed3)

par(mfrow = c(1,1))
plot(model_mixed3)
qqnorm(resid(model_mixed3))
qqline(resid(model_mixed3))
```


  
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
