---
title: "Simple and Mixed Model Analysis"
author: "Alexis Swanson"
date: "3/3/2021"
output:
  word_document: default
  html_document: default
---

#Histograms of the 3 Approaches for tc
  
```{r, echo=FALSE, message=FALSE}

data <- read.csv("/Users/swans/Documents/Research/SiteData_Mar_3_2021.csv")

hist(data$tc.B)
hist(data$tc.S)
hist(data$tc.I)

hist(data$D50)


hist(log(data$tc.B))
data$log.tc = log(data$tc.B)
data$log.tc[data$log.tc== "-Inf"] <- 0

range(data$tc.B)
range(data$tc.S)
range(data$tc.I)

boxplot(log.tc ~ Test.Location, data=data)
boxplot(log.tc ~ Land.Use, data=data)
boxplot(log.tc ~ Soil.Texture, data=data)
```

The ranges are for Blaisdell, Scour Depth and the Iterative solutions respectively. 

After our discussion on 2/18 the concensus was to move forward using the Blaisdell solution.

# First Linear Model
  # All Predictors

```{r, echo=FALSE, message=FALSE}

library(ggplot2)

model_simple <-lm(tc.B ~ Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_simple)

par(mfrow = c(2,2))
plot(model_simple)
par(mfrow = c(1,1))
qqnorm(resid(model_simple))
qqline(resid(model_simple))


library(tidyverse)

data$Fitted <- model_simple$fitted.values

# Create a plot showing a one to one of the simulated and observed tc values
ggplot() + 
  geom_point(data = data, 
             aes(x = Fitted, y = log.tc),
             color = "black") +
  geom_abline(intercept=0, slope=1) +
  theme_bw() + xlim(-5,5) + ylim(-5,5) +
  labs(x = expression("Simulated"),
       y = expression(" Observed"))
```

Here I also plotted box plots to see if there was a significant difference in the categorical variables. The field and lab almost show a statistical difference but are still overlapping slightly. The rural, urban and ultra urban are not showing much of a difference. The soil texture however seems to have something going on where the critical shear stress could be statiscially different for each soil texture.



# Mixed Model with 3 Random Factors
  # Test Location and Land Use
  
```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed <-lmer(log.tc~(1|Test.Location) + (1|Land.Use) + (1|Soil.Texture) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed)
anova(model_mixed)
ranef(model_mixed)

par(mfrow = c(1,1))
plot(model_mixed)
qqnorm(resid(model_mixed))
qqline(resid(model_mixed))

```


# Mixed Model with only Test Location as Random Factor

```{r, echo=FALSE, message=FALSE}

model_mixed2 <-lmer(log.tc~(1|Test.Location) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed2)
anova(model_mixed2)
ranef(model_mixed2)

par(mfrow = c(1,1))
plot(model_mixed2)
qqnorm(resid(model_mixed2))
qqline(resid(model_mixed2))

```


# Mixed Model with only Land Use as Random Factor

```{r, echo=FALSE, message=FALSE}

model_mixed3 <-lmer(tc.I~(1|Land.Use) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed3)
anova(model_mixed3)
ranef(model_mixed3)

par(mfrow = c(1,1))
plot(model_mixed3)
qqnorm(resid(model_mixed3))
qqline(resid(model_mixed3))
```

# Mixed Model with only Soil Texture as Random Factor

```{r, echo=FALSE, message=FALSE}

model_mixed4 <-lmer(tc.I~(1|Soil.Texture) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed4)
anova(model_mixed4)
ranef(model_mixed4)

par(mfrow = c(1,1))
plot(model_mixed4)
qqnorm(resid(model_mixed4))
qqline(resid(model_mixed4))
```

# Mixed Model with Soil Texture and Land Use as Random Factors

```{r, echo=FALSE, message=FALSE}

model_mixed5 <-lmer(tc.B~(1|Soil.Texture) + (1|Land.Use) + Temp + EC + Cor.Moist + D50 + Cu + Cc + GSD + Pass200, data = data)
summary(model_mixed5)
anova(model_mixed5)
ranef(model_mixed5)

par(mfrow = c(1,1))
plot(model_mixed5)
qqnorm(resid(model_mixed5))
qqline(resid(model_mixed5))

library(tidyverse)

data$Fitted_Mixed <- predict(model_mixed5)

data$Residuals_Mixed <- resid(model_mixed5)
data$Residuals_Simple <- resid(model_simple)

#Create a plot showing both residuals
ggplot() + 
  geom_point(data = data, 
             aes(x = Residuals_Mixed, y = log.tc),
             color = "red") +
  geom_point(data = data,
             aes(x = Residuals_Simple, y = log.tc),
             color = "black") +
  geom_abline(intercept=0, slope=1) +
  theme_bw() + xlim(-5,5) + ylim(-5,5) +
  labs(x = expression("Simulated"),
       y = expression(" Observed"))

# Create a plot showing a one to one of the simulated and observed tc values
ggplot() + 
  geom_point(data = data, 
             aes(x = Fitted_Mixed, y = tc.B),
             color = "red") +
  geom_point(data = data,
             aes(x = Fitted, y = tc.B),
             color = "black") +
  geom_abline(intercept=0, slope=1) +
  theme_bw() + xlim(-2.5,5) + ylim(-2.5,5) +
  labs(x = expression("Simulated"),
       y = expression(" Observed"))
```


# Conclusion


# Load new data that includes bulk density and water temperature (59 observations)
``````{r, echo=FALSE, message=FALSE}

data_limited <- read.csv("/Users/swans/Documents/Research/SiteData_Mar_3_2021_limited.csv")

data_limited$log.tc = log(data_limited$tc.B)
data_limited$log.tc[data_limited$log.tc== "-Inf"] <- 0

```

# Simple Linear Model using limited data with all predictors

```{r, echo=FALSE, message=FALSE}

library(ggplot2)

model_simple_l <-lm(log.tc ~ Temp + EC + Cor.Moist + WTemp + BD + D50 + Cu + Cc + GSD + Pass200, data = data_limited)
summary(model_simple_l)

par(mfrow = c(2,2))
plot(model_simple_l)
par(mfrow = c(1,1))
qqnorm(resid(model_simple_l))
qqline(resid(model_simple_l))
```

#Simple linear model using limited data and only signifcant predictors

```{r, echo=FALSE, message=FALSE}

library(ggplot2)

model_simple_l2 <-lm(log.tc ~ WTemp + D50 + Cu, data = data_limited)
summary(model_simple_l2)

par(mfrow = c(2,2))
plot(model_simple_l2)
par(mfrow = c(1,1))
qqnorm(resid(model_simple_l2))
qqline(resid(model_simple_l2))


```

# Mixed Model with limited data and only Land Use as a random factor
```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed_l1 <-lmer(tc.I~(1|Land.Use) + Temp + EC + Cor.Moist + WTemp + BD + D50 + Cu + Cc + GSD + Pass200, data = data_limited)
summary(model_mixed_l1)
anova(model_mixed_l1)
ranef(model_mixed_l1)

par(mfrow = c(1,1))
plot(model_mixed_l1)
qqnorm(resid(model_mixed_l1))
qqline(resid(model_mixed_l1))
```


# Mixed Model with limited data and only Soil Texture as a random factor
```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed_l2 <-lmer(tc.I~(1|Soil.Texture) + Temp + EC + Cor.Moist + WTemp + BD + D50 + Cu + Cc + GSD + Pass200, data = data_limited)
summary(model_mixed_l2)
anova(model_mixed_l2)
ranef(model_mixed_l2)

par(mfrow = c(1,1))
plot(model_mixed_l2)
qqnorm(resid(model_mixed_l2))
qqline(resid(model_mixed_l2))
```

# Mixed Model with limited data and both Land Use and Soil Texture as random factors

```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed_l3 <-lmer(tc.I~(1|Soil.Texture) + (1|Land.Use) + Temp + EC + Cor.Moist + WTemp + BD + D50 + Cu + Cc + GSD + Pass200, data = data_limited)
summary(model_mixed_l3)
anova(model_mixed_l3)
ranef(model_mixed_l3)

par(mfrow = c(1,1))
plot(model_mixed_l3)
qqnorm(resid(model_mixed_l3))
qqline(resid(model_mixed_l3))
```

# Mixed Model with limited data, Soil Texture and only sig variables from simple linear model

```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed_l4 <-lmer(tc.I~(1|Soil.Texture) + WTemp + D50 + Cu, data = data_limited)
summary(model_mixed_l4)
anova(model_mixed_l4)
ranef(model_mixed_l4)

par(mfrow = c(1,1))
plot(model_mixed_l4)
qqnorm(resid(model_mixed_l4))
qqline(resid(model_mixed_l4))
```

# Load limited data based on water temp, this dataset does not include BD (69 observations)

``````{r, echo=FALSE, message=FALSE}

data_wtemp <- read.csv("/Users/swans/Documents/Research/SiteData_Mar_3_2021_wtemp.csv")

data_wtemp$log.tc = log(data_wtemp$tc.B)
data_wtemp$log.tc[data_wtemp$log.tc== "-Inf"] <- 0

```

# Simple Linear Model using limited data based on watertemp with all predictors

```{r, echo=FALSE, message=FALSE}

library(ggplot2)

model_simple_w <-lm(tc.B ~ Temp + EC + Cor.Moist + WTemp + D50 + Cu + Cc + GSD + Pass200, data = data_wtemp)
summary(model_simple_w)

par(mfrow = c(2,2))
plot(model_simple_w)
par(mfrow = c(1,1))
qqnorm(resid(model_simple_w))
qqline(resid(model_simple_w))

library(tidyverse)

data_wtemp$Fitted <- model_simple_w$fitted.values

# Create a plot showing a one to one of the simulated and observed tc values
ggplot() + 
  geom_point(data = data_wtemp, 
             aes(x = Fitted, y = tc.B),
             color = "black") +
  geom_abline(intercept=0, slope=1) +
  theme_bw() + xlim(-5,5) + ylim(-5,5) +
  labs(x = expression("Simulated"),
       y = expression(" Observed"))

# Create a plot showing a one to one of the simulated and observed tc values for the three models
ggplot() + 
  geom_point(data = data, 
             aes(x = Fitted_Mixed, y = tc.B),
             color = "red") +
  geom_point(data = data,
             aes(x = Fitted, y = tc.B),
             color = "black") +
  geom_point(data = data_wtemp,
             aes(x = Fitted, y = tc.B),
             color = "blue") +
  geom_abline(intercept=0, slope=1) +
  theme_bw() + xlim(-5,5) + ylim(-5,5) +
  labs(x = expression("Simulated"),
       y = expression(" Observed"))
```


# Mixed Model with limited data and only Land Use as a random factor
```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed_w1 <-lmer(tc.I~(1|Land.Use) + Temp + EC + Cor.Moist + WTemp + D50 + Cu + Cc + GSD + Pass200, data = data_wtemp)
summary(model_mixed_w1)
anova(model_mixed_w1)
ranef(model_mixed_w1)

par(mfrow = c(1,1))
plot(model_mixed_w1)
qqnorm(resid(model_mixed_w1))
qqline(resid(model_mixed_w1))
```


# Mixed Model with limited data and only Soil Texture as a random factor
```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed_w2 <-lmer(tc.B~(1|Soil.Texture) + Temp + EC + Cor.Moist + WTemp + D50 + Cu + Cc + GSD + Pass200, data = data_wtemp)
summary(model_mixed_w2)
anova(model_mixed_w2)
ranef(model_mixed_w2)

par(mfrow = c(1,1))
plot(model_mixed_w2)
qqnorm(resid(model_mixed_w2))
qqline(resid(model_mixed_w2))
```

# Mixed Model with limited data and both Land Use and Soil Texture as random factors

```{r, echo=FALSE, message=FALSE}

library(lme4)

model_mixed_w3 <-lmer(tc.B~(1|Soil.Texture) + (1|Land.Use) + Temp + EC + Cor.Moist + WTemp + D50 + Cu + Cc + GSD + Pass200, data = data_wtemp)
summary(model_mixed_w3)
anova(model_mixed_w3)
ranef(model_mixed_w3)

par(mfrow = c(1,1))
plot(model_mixed_w3)
qqnorm(resid(model_mixed_w3))
qqline(resid(model_mixed_w3))

data_wtemp$Fitted_Mixed <- predict(model_mixed_w3)


```

# Calculate Model Statistics for the Four Models and Compare

```{r}

# Create a plot showing a one to one of the simulated and observed tc values for the four models
ggplot() + 
  geom_point(data = data, 
             aes(x = Fitted_Mixed, y = tc.B),
             color = "red") +
  geom_point(data = data,
             aes(x = Fitted, y = tc.B),
             color = "black") +
  geom_point(data = data_wtemp,
             aes(x = Fitted, y = tc.B),
             color = "blue") +
  geom_point(data = data_wtemp,
             aes(x = Fitted_Mixed, y = tc.B),
             color = "green") +
  geom_abline(intercept=0, slope=1) +
  theme_bw() + xlim(0,5) + ylim(0,5) +
  labs(x = expression("Simulated"),
       y = expression(" Observed"))

# Set up an empty data frame to compile results of the four models
model_statistics <- data.frame(matrix(ncol = 5, nrow = 4))
x <- c("Model", "RMSE", "EF", "r", "Bias")
colnames(model_statistics) <- x

model_statistics$Model <- c("Simple Model", "Mixed Model", "Simple with WTemp", "Mixed with WTemp")

#Calculate the RMSE
RMSE_Simple <- sqrt(mean((data$tc.B - data$Fitted)^2))
RMSE_Mixed <- sqrt(mean((data$tc.B - data$Fitted_Mixed)^2))
RMSE_WTemp_Simple <- sqrt(mean((data$tc.B - data_wtemp$Fitted)^2))
RMSE_WTemp_Mixed <- sqrt(mean((data$tc.B - data_wtemp$Fitted_Mixed)^2))

model_statistics$RMSE <- c(RMSE_Simple, RMSE_Mixed, RMSE_WTemp_Simple, RMSE_WTemp_Mixed)

#Calculate EF
EF_Simple <- 1 - (sum((data$tc.B-data$Fitted)^2)/sum((data$tc.B-mean(data$Fitted))^2))
EF_Mixed <- 1 - (sum((data$tc.B-data$Fitted_Mixed)^2)/sum((data$tc.B-mean(data$Fitted_Mixed))^2))
EF_WTemp_Simple <- 1 - (sum((data_wtemp$tc.B-data_wtemp$Fitted)^2)/sum((data_wtemp$tc.B-mean(data_wtemp$Fitted))^2))
EF_WTemp_Mixed <- 1 - (sum((data_wtemp$tc.B-data_wtemp$Fitted_Mixed)^2)/sum((data_wtemp$tc.B-mean(data_wtemp$Fitted_Mixed))^2))

model_statistics$EF <- c(EF_Simple, EF_Mixed, EF_WTemp_Simple, EF_WTemp_Mixed)

#Calculate r
r_Simple <- cor(data$tc.B,data$Fitted)
r_Mixed <- cor(data$tc.B,data$Fitted_Mixed)
r_WTemp_Simple <- cor(data_wtemp$tc.B,data_wtemp$Fitted)
r_WTemp_Mixed <- cor(data_wtemp$tc.B,data_wtemp$Fitted_Mixed)

model_statistics$r <- c(r_Simple,r_Mixed, r_WTemp_Simple, r_WTemp_Mixed)

#Calculate bias
Bias_Simple <- mean(data$tc.B-data$Fitted)
Bias_Mixed <- mean(data$tc.B-data$Fitted_Mixed)
Bias_WTemp_Simple <- mean(data_wtemp$tc.B-data_wtemp$Fitted)
Bias_WTemp_Mixed <- mean(data_wtemp$tc.B-data_wtemp$Fitted_Mixed)

model_statistics$Bias <- c(Bias_Simple, Bias_Mixed, Bias_WTemp_Simple, Bias_WTemp_Mixed)

model_statistics


```

  
*This is only for my reference when writing the code in R*  
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
